### 2020-07-10 steadystate.tex
Katy comment edits

With the verification of Moltres' neutronics modeling capabilities in the
context of the \gls{MSFR}, this chapter will cover the steady-state
multiphysics simulation results from Moltres.

The procedure for obtaining the steady-state results involved several steps
due to the tightly coupled \glspl{PDE}. First, a preliminary transient
simulation of fluid flow in the \gls{MSFR} core was run, starting from zero
inlet velocity and gradually ramping up to match the nominal flow rate (4.5
m$^3$ s$^{-1}$); otherwise Moltres had difficulty converging to the desired
fully developed flow profile.
Next, these fully developed flow values were imported as initial values for
velocity in the full time-dependent simulation modeling the full coupled
neutronics and thermal-hydraulics multiphysics model. The initial values for
the temperature and neutron group flux distributions are 953 K and
$1 \times 10^{14}$ cm$^{-2}$ s$^{-1}$ uniformly throughout the geometry.
Finally, this work assumes that steady
state is reached when the volume integral values of every variable remain
constant (up to 6 sig. fig.) for at least four seconds in the simulation; this
time period corresponds to the nominal circulation time of the \gls{MSFR}.

For a direct comparison with the steady-state results from the Polimi and
TUDelft models \cite{aufiero_development_2014}, this chapter will first
present the steady-state results without modeling decay heat. After this
comparison, the chapter separately discusses the minor differences in the
results from decay heat modeling in the last subsection.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{.365\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models. The colors represent
    temperature according to the respective colorbars and the arrows
    represent velocity fields.}
    \label{fig:flow-temp}
\end{figure}

\begin{figure}[t!]
    \centering
    \includegraphics[width=.6\textwidth]{flow}
    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
    The colors represent velocity magnitude according to the colorbar on the
    right.}
    \label{fig:flow}
\end{figure}

Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
models. Figure \ref{fig:flow} provides an alternate view of the flow profile
through flow streamlines superimposed on the velocity magnitude distribution.
The results from Moltres show good qualitative agreement with the Polimi and
TUDelft models \cite{fiorina_modelling_2014}; the plots show similar
flow and hotspot features in all three models. Furthermore, the highest salt
velocities in all three models occur at the inlet, outlet, and at core
half-height approximately 0.40 m away from the central axis. A large
recirculation region forms near the blanket tank walls from turbulent flow.
Inertial forces dominate over viscous forces to form this large eddy. The main
difference between Moltres and Fiorina et al.'s models is the flow profile
near the central axis at the top and bottom of the core.
The Polimi and TUDelft models predict relatively stagnant flow in these
regions without recirculation. Moltres, on the other hand, predicts explicit
recirculating flow in these regions. This is due to our constant turbulent
viscosity approximation in Moltres. The k-$\epsilon$ turbulence models in the
Polimi and TUDelft models predict that the turbulent viscosity in these
regions is as high as 100 Pa s, much higher than our 40 Pa s approximation.

@@

Although the temperatures at the hotspots are well below the melting point of the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
blanket tank structure and induce relatively faster salt corrosion rates. A
sudden, large reactivity insertion could push fuel salt temperatures above the
melting point of the Ni-alloy and cause irreversible damage. Furthermore,
the reservoir of hot fuel salt may cause unpredictable behavior during
transient scenarios when the flow profile undergoes a drastic change.
Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
hourglass-shaped design to optimize flow distribution and prevent these
recirculation zones and hotspots from forming. A study of this new design
using Moltres is a potential subject for future work when a proper turbulence
model is in place.

@@

The neutron flux distribution represents the heat source distribution in a
nuclear reactor. Figure \ref{fig:neutronflux} shows the neutron flux
distributions in the core for all six neutron energy groups, and Figure
\ref{fig:axialradial} shows the axial and radial fluxes along the center of
the core and at reactor half-height, respectively.
The distributions are highly symmetric along the
central and horizontal axes, as expected of a cylindrical reactor design. The
relatively lower temperatures near the center of the core promotes the neutron
flux peaking but it is of little concern as no structural parts vulnerable to
neutron damage exist in that region. The peak total flux at the center is
$9.80 \times 10^{15}$ cm$^{-2}\cdot$s$^{-1}$, which is close to values
reported by Fiorina et al. \cite{fiorina_molten_2013} and Aufiero et al.
\cite{aufiero_development_2014} as shown in Table \ref{table:peak-flux}. The
peak flux value from this paper is slightly higher as this work used the
steady-state temperature distribution while Fiorina et al. and Aufiero et al.
imposed a uniform temperature distribution at 973 K.

@@

\clearpage

>>>>

Building on Chapter \ref{chap:nts} which verified Moltres' neutronics modeling
capabilities in the context of the \gls{MSFR}, this chapter will cover the
steady-state multiphysics simulation results of the \gls{MSFR} using Moltres.
The steady-state results depend heavily on both neutronics and
thermal-hydraulics solving capabilities in Moltres. Thus, this exercise is the
first verification step for Moltres as a multiphysics \gls{MSR} simulation
tool. This chapter will specifically discuss the temperature, velocity,
neutron flux, and precursor distributions and compare them with data from the
literature. The steady-state operating conditions will also serve as the
initial conditions for the subsequent accident transient simulations.

This chapter will first
present a summary of the procedure for obtaining the steady-state results.
Next, the chapter discusses the steady-state results without modeling decay
heat in direct comparison with the steady-state results from the Polimi and
TUDelft models \cite{aufiero_development_2014}. After this
comparison, the chapter separately discusses the minor differences in the
results from decay heat modeling in the last subsection.

\section{Simulation Procedure}

The procedure for obtaining the results for the steady-state operating
conditions involved several steps
due to the tightly coupled \glspl{PDE}. First, a preliminary transient
simulation of fluid flow in the \gls{MSFR} core was run, starting from zero
inlet velocity and gradually ramping up to match the nominal flow rate (4.5
m$^3$ s$^{-1}$); otherwise Moltres had difficulty converging to the desired
fully developed flow profile.
Next, these fully developed flow values were imported as initial values for
velocity in the full time-dependent simulation modeling the full coupled
neutronics and thermal-hydraulics multiphysics model. The initial values for
the temperature and neutron group flux distributions are 953 K and
$1 \times 10^{14}$ cm$^{-2}$ s$^{-1}$ uniformly throughout the geometry.
Finally, this work assumes that steady state is reached when the volume
integral values of every variable remain constant (up to 6 significant
figures) for at least four seconds in the simulation; this
time period corresponds to the nominal circulation time of the \gls{MSFR}.

@@

Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
models. Figure \ref{fig:flow} provides an alternate view of the flow profile
through flow streamlines superimposed on the velocity magnitude distribution.
The results from Moltres show good qualitative agreement with the Polimi and
TUDelft models \cite{fiorina_modelling_2014}; the plots show similar
flow and hotspot features in all three models. Furthermore, the highest salt
velocities in all three models occur at the inlet, outlet, and at core
half-height approximately 0.40 m away from the central axis. A large
recirculation region forms near the blanket tank walls from turbulent flow.
Inertial forces dominate over viscous forces to form this large eddy. The main
difference between Moltres and Fiorina et al.'s models is the flow profile
near the central axis at the top and bottom of the core.
The Polimi and TUDelft models predict relatively stagnant flow in these
regions without recirculation. Moltres, on the other hand, predicts explicit
recirculating flow in these regions. This is due to our constant turbulent
viscosity approximation in Moltres. The k-$\epsilon$ turbulence models in the
Polimi and TUDelft models predict that the turbulent viscosity in these
regions is as high as 100 Pa$\cdot$s, much higher than our 40 Pa$\cdot$s
approximation.

\begin{figure}[htbp!]
    \centering
    \begin{subfigure}[t]{.365\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models. The colors represent
    temperature according to the respective color bars and the arrows
    represent velocity fields.}
    \label{fig:flow-temp}
\end{figure}

\begin{figure}[htbp!]
    \centering
    \includegraphics[width=.6\textwidth]{flow}
    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
    The colors represent velocity magnitude according to the color bar on the
    right.}
    \label{fig:flow}
\end{figure}

@@

Although the temperatures at the hotspots are well below the melting point of the Ni-alloy structure (1500 K), they may cause undue thermal stress on the
blanket tank structure and accelerate salt corrosion rates. A
sudden, large reactivity insertion could push fuel salt temperatures above the
melting point of the Ni-alloy and cause irreversible damage. Furthermore,
the reservoir of hot fuel salt may cause unpredictable behavior during
transient scenarios when the flow profile fluctuates.
Thus, Rouch et al. \cite{rouch_preliminary_2014} developed an improved
hourglass-shaped design for the central core region to optimize flow
distribution and prevent these
recirculation zones and hotspots from forming. A study of this new design
using Moltres is a potential subject for future work when a proper turbulence
model is in place.

@@

The neutron flux distribution represents the heat source distribution in a
nuclear reactor. Figure \ref{fig:neutronflux} shows the neutron flux
distributions in the core for all six neutron energy groups, and Figure
\ref{fig:axialradial} shows the axial and radial fluxes along the center of
the core and at reactor half-height, respectively.
The distributions are highly symmetric along the
central and horizontal axes, as expected of a cylindrical reactor design. The
lower temperatures near the center of the core promote the neutron
flux peaking but it is of little concern as no neutronically vulnerable
structures exist in that region. The peak total flux at the center is
$9.80 \times 10^{15}$ cm$^{-2}\cdot$s$^{-1}$, which is close to values
reported by Fiorina et al. \cite{fiorina_molten_2013} and Aufiero et al.
\cite{aufiero_development_2014} as shown in Table \ref{table:peak-flux}. The
peak flux value from this paper is slightly higher as this work used the
heterogeneous temperature distribution in Figure \ref{fig:flow-temp} while
Fiorina et al. and Aufiero et al. imposed uniform temperature distributions
at 973 K.

@@
